Temporal graph realization from fastest paths

### 2024-01-13 SAND2024/SAND24-longversion.tex
Fixed proof for tree vertices of the FPT algorithm

From the construction of $Z$ (\ie by exhaustively removing vertices of degree one from~$G$), it follows that $G[Z]$ (graph induced in $G$ by $Z$) consists of disjoint trees. 
Each of these trees has a unique neighbor $u_i$ in $G'$. Denote by $T_i$ the tree that is obtained by such a vertex $u_i$ and all the trees from $G[Z]$ that are incident to $u_i$. Then  we call $u_i$ the \emph{clip vertex} of the tree $T_i$. 
For each of these trees~$T_{i}$, we select an arbitrary vertex $r_i \in Z$ that is a neighbor of the clip vertex $u_i$ in $T_i$,
and call it the \emph{representative vertex} of the tree $T_{u_i}$. 
We now define by $Z^*$ the set of all representatives $r_i$ of the trees $T_i$,
where the clip vertex $u_i$ of $T_i$ is a vertex of interest, \ie
\[
Z^* = \{ r_i  :  r_i \in Z \text{ is a neighbor of the clip vertex $u_i$ in the tree $T_i$, and $u_i \in U$}\}.
\]
Since there are $O(k)$ vertices of interest, we get that $|Z^*| \in O(k)$.
Finally, the set of \emph{important vertices} is defined as the set $U \cup U^{\ast} \cup Z^{\ast}$.
For an illustration see \cref{fig:labelingVertices}. Note that determining sets $U$, $U^*$, and $Z^*$ takes linear time.

@@

%Let us remember the following basic property of the problem \deltaExact.
Recall that a labeling $\lambda$ of $G$ satisfies $D$ if the duration of a fastest temporal path from vertex $v_i$ to $v_j$ equals $D_{v_i, v_j}$.
%, and no other temporal path from $v_i$ to $v_j$ is faster.
In order to find a labeling that satisfies this property we split our analysis in nine cases.
We consider fastest temporal paths 
where the starting vertex is in one of the sets $U,V(G') \setminus U, Z$,
and similarly the destination vertex is in one of the sets $U,V(G') \setminus U, Z$.
In each of these cases we guess the underlying path $P$ that at least one fastest temporal path from the vertex $v_i$ to $v_j$ follows, 
which results in one equality constraint for the labels on the path $P$. 
For all other temporal paths from $v_i$ to $v_j$ we know that they cannot be faster, so we introduce inequality constraints for them.
This results in producing $f(k) \cdot |D| ^ {O(1)}$ constraints. 
Note that we have to do this while keeping the total number of variables upper-bounded by some function in $k$.
%We have to guarantee only that the number of guesses of fastest temporal paths is exponential in $k$ and not in $n$.

For an easier understanding and the analysis of the algorithm we give the following definition.
\begin{definition}\label{def:segments}
    Let $U \subseteq V(G')$ be a set of vertices of interest and let $u,v \in U$.
    A path $P = (u=v_1,v_2, \dots, v_p = v)$  in graph $G'$, where all inner vertices are not in $U$, \ie $v_i \notin U$ for all $i \in \{ 2, 3, \dots, p-1\}$,
    is called a \emph{segment} from $u$ to $v$. We denote it as $S_{u,v}$.
\end{definition}
%From the definition it follows, that there are $O(k^2)$ segments in $G'$. \todo{True but does not follow directly just from the definition}
Note from \cref{def:segments} we get that $S_{u,v} \neq S_{v,u}$, since we consider paths to be directed. Observe that a temporal path in $G'$ between two vertices of interest is either a segment, or consists of a sequence of some segments. Furthermore, since we have at most $4k$ interesting vertices in $G'$, we can deduce the following important result.
\begin{corollary}\label{obs:FPT-k2segments}
    There are at most $O(k^2)$ segments in $G'$.
\end{corollary}
%In the analysis of the algorithm we rely heavily on the fact that there are $O(k^2)$ segments in $G'$.
%, which we prove later on (see~\cref{obs:FPT-k2segments}).
%For a pair $u,v$ of vertices in $U$, there are $O(k!)=O(k^k)$ possible paths in $G'$ between them.

@@

\begin{lemma}\label{lemma:FPT-unlabeldPaths-01}
    Suppose that
    $S_{u,v}, S_{w,z}$ are two segments with $ v_i \in S_{u,v}$ and $z_j \in S_{w,z}$,
    where $z_j$ is a split vertex of $v_i$ in the segment $S_{w,z}$.
    W.l.o.g. suppose that the fastest temporal path from $v_i$ to $z_j$ travels through vertices $u$ and $w$.
    Then the fastest temporal path from $v_i$ to any other vertex of $S_{w,z}$, that is closer to $w$,
    travels through the same two vertices $u$ and $w$.
    Similarly it holds for the cases when the fastest temporal path travels through $w,v$ or $z,u$ or~$z,v$.
\end{lemma}
\begin{proof}
    Let $z_\ell$ be a vertex of $S_{w,z}$, that is closer to $w$ than $z$ in the segment.
    Let us denote with $P_{v_i,z_j}$ the underlying path of the fastest temporal path from $v_i$ to $z_j$.
    Denote with $P_{v_i,z_j}^\ell$
    the subpath of the fastest temporal path from $v_i$ to $z_j$, that terminates in $z_\ell$.
    We want to show that $P_{v_i,z_j}^\ell$ is an underlying path of a fastest temporal path from $v_i$ to $z_j$.
    Let us observe the following possibilities.

@@

    Third, suppose that the fastest temporal path from $v_i$ to $z_\ell$ travels through vertices $v$ and $z$.
    Denote this path as $P^3_{v_i,z_\ell}$.
    Then the duration of the temporal path from $v_i$ to $z_j$ using the subpath of $P^3_{v_i,z_\ell}$, would be strictly smaller than the duration of $(P_{v_i,z_j},\lambda)$, which cannot be possible.
\end{proof}

\begin{lemma}\label{lemma:FPT-noUndeterminedEdgeInSegment}
Let $S_{u,v}$ be a segment in $G$ of length at least $5$, denoted as $S_{u,v}= (u=v_1,v_2, \dots, v_p = v)$, where $p > 5$.
It is impossible for an inner edge $f = v_i v_{i+1}$ from $S_{u,v} \setminus \{u,v\}$ (where $f$ is an edge that is not incident to a vertex from $U$)
to not be a part of any fastest temporal path, of length at least $2$ between vertices in $S_{u,v}$.
In other words, there must exist a pair $v_j, v_{j'} \in S_{u,v}$ \st the fastest temporal path from $v_j$ to $v_{j'}$ passes through $f$.
In the case where $p = 5$ all temporal paths of length $2$ avoid $f$ if and only if $f$ has the same label as both of the edges incident to it, while the label of the last remaining edge is a function of $\lambda(f)$.
\end{lemma}

\begin{proof}
    For an easier understanding and better readability we present the proof for $S_{u,v}$ of fixed length $5$.
    The case where $S_{u,v}$ is longer easily follows from presented results.

@@

    \emph{Case 1: }$f = v_2v_3$.
    Let us observe the case of fastest temporal paths between $v_{1}$ and $v_{3}$.
    Denote with $Q = (v_{1}, v_2, v_3)$ and with $P' = (v_3,v_4,v_5,v_6)$.
    From our proposition it follows that
    \begin{itemize}
        \item the fastest temporal path $P ^ +$ from $v_1$ to $v_3$ 
    is of the following form 
    $P^+ = v_{1}  \leadsto v_6 \rightarrow v_5 \rightarrow v_4 \rightarrow v_3$,
    and
        \item the fastest temporal path $P ^ -$ from $v_{3}$ to $v_{1}$ 
    is of the following form 
    $P^- = v_{3} \rightarrow v_{4} \rightarrow v_5 \rightarrow v_6 \leadsto v_1$.
    \end{itemize}
    It follows that 
    $
    d(v_{1}, v_{3}, P^+) \leq d(v_{1}, v_{3}, Q),
    $
    and
    $
    d(v_{1}, v_{3}, P^-) \leq d(v_{1}, v_{3}, Q)
    $. 
    Note that $d(v_{1}, v_{3}, P^+) \geq 1 + d(v_6,v_3,P')$,
    and by the definition $d(v_6,v_3,P') = 1 + (t_4 - t_5)_\Delta + (t_3 - t_4)_\Delta$,
    where $(t_i - t_j)_\Delta$ denotes the difference of two consecutive labels $t_i, t_j$ modulo $\Delta$.
    Similarly holds for $d(v_{1}, v_{3}, P^+)$.
    Summing now both of the above equations we get
    \begin{equation}
    \begin{split} \label{eq:FPT-prf-fUnlabeled}
        d(v_{1}, v_{3}, P^+) + d(v_{3}, v_{1}, P^-) &\leq 
        d(v_{1}, v_{3}, Q) + d(v_{3}, v_{1}, Q) \\
        %
        1 + d(v_6,v_3,P') + 1 + d(v_3,v_6,P') &\leq d(v_{1}, v_{3}, Q) + d(v_{3}, v_{1}, Q) \\
        %
        3 + (t_4 - t_5)_\Delta + (t_3 - t_4)_\Delta +
        1 + (t_4 - t_3)_\Delta + (t_5 - t_4)_\Delta 
        &\leq 
        1 + (t_2 - t_1)_\Delta + 
        1 + (t_1 - t_2)_\Delta\\
        %
        (t_4 - t_5)_\Delta + (t_5 - t_4)_\Delta +
        (t_4 - t_3)_\Delta + (t_3 - t_4)_\Delta  + 2 
        &\leq 
        (t_2 - t_1)_\Delta + (t_1 - t_2)_\Delta.
    \end{split}
    \end{equation}
    Note that if $t_i \neq t_j$ we get that 
    the sum
    $(t_i - t_j)_\Delta + (t_j - t_i)_\Delta$ equals exactly $\Delta$,
    and if $t_i = t_j$ the sum equals $2\Delta$.
    This follows from the definition of travel delays at vertices (see \cref{obs:travel-delays-both-directions}).
    Therefore we get from \cref{eq:FPT-prf-fUnlabeled}, 
    that the right part is at most $2 \Delta$, while the left part is at least $2 \Delta + 1$,
    for any relation of labels $t_1,t_2, \dots, t_5$,
    which is a contradiction.

@@

    Let us observe now the case when $S_{u,v} = (u=v_1,v_2, v_3, v_4, v_5=v)$ is of length $4$.
    Let $f = v_2 v_3$ (the case with $f = v_3 v_4$ is symmetrical).
    Suppose that the fastest temporal paths between $v_1$ and $v_3$ do not use the edge $f$.
    We denote with $R^+$ the fastest path from $v_1$ to $v_3$, 
    which is of the form $u \leadsto v \rightarrow v_4 \rightarrow v_3$,
    and similarly 
    with $R^-$ the fastest path from $v_3$ to $v_1$, which is
    of the form $v_3 \rightarrow v_4 \rightarrow v \leadsto u$.
    We denote with $R' = (v_3, v_4, v_5)$ and with $S = (v_1,v_2,v_3)$.
    Again we get the following.
    \begin{equation*}
    \begin{split}
        d(v_{1}, v_{3}, R^+) + d(v_{3}, v_{1}, R^-) &\leq 
        d(v_{1}, v_{3}, S) + d(v_{3}, v_{1}, S) \\
        %
        1 + d(v_5,v_3,R') + 1 + d(v_3,v_5,R') &\leq d(v_{1}, v_{3}, S) + d(v_{3}, v_{1}, S) \\
        %
        (t_3 - t_4)_\Delta + (t_4 - t_3)_\Delta + 2 
        &\leq 
        (t_2 - t_1)_\Delta + (t_1 - t_2)_\Delta.        
    \end{split}
    \end{equation*}
    The only case when the equation has a valid solution is when $t_1 = t_2$ and $t_3 \neq t_4$,
    as in this case the left hand side evaluates to $\Delta + 2$, while the right side evaluates to $2 \Delta$.
    %
    Repeating the analysis for the fastest paths between $v_2$ and $v_4$,
    we conclude that the only valid solution is when $t_2 = t_3$ and $t_1 \neq t_4$.
    %
    Altogether, we get that $f$ is not a part of any fastest path of length $2$ in $S_{u,v}$ if and only if the label of edge $f$ is the same as the labels on the edges incident to it, while the last remaining edge has a different label.
    %
    Note now that the fastest temporal path from $v_2$ to $v_4$ must first use the edge $u v_2$ and finish with the edge $v_4 v_5$, and it has to be of duration $D_{v_2, v_4}$. Using \cref{lemma:FPT-uv-Labelalledges} we determine the label of the edge $v_4v_5$ as a function of $\lambda(f)$.
\end{proof}

@@

Each specific calculation takes linear time, since there are $O(k^2)$ segments, this calculation takes $O(k^4)$ time.

@@

the edges of every segment are fully determined, 

@@

For this we use the fact that at least one of the temporal paths between vertices in $S_{u,v}$ has to pass through $f$, when $S_{u,v}$ has at least $5$ edges (see~\cref{lemma:FPT-noUndeterminedEdgeInSegment}).

@@

    Thus we create $4$ guesses for every such segment $S_{u,v}$, and $O(k^2)$ new guesses in total, as there are at most $O(k^2)$ segments.

@@

Once we know which two fastest temporal paths pass through $f$, 

@@

Note, from \cref{lemma:FPT-noUndeterminedEdgeInSegment} it follows that the above procedure holds only for segments with at least $5$ edges.
In the case when a segment has $4$ edges, it can happen that the fastest temporal paths from above, do not traverse $f$. But in this case, we know that three edges have the same label, denote it $c$, while the label of the last one is a function of $c$.

@@

At this point $G'$ is a graph, where each edge $e$ has a value for its label $\lambda(e)$
that depends on (\ie is a function of) some other label $\lambda(f)$ of edge $f$,
or it depends on no other label.
We now describe how we create variables and start building our ILP instances. 
For every edge $e$ in $G'$ that is incident to a vertex of interest, we create a variable $x_e$ that can have values from $ \{1, 2, \dots, \Delta\}$.
Besides that, we create one variable for each edge that is still not yet determined on a segment.
Since each vertex of interest is incident to at most $k$ edges, and each segment has at most one extra not yet determined edge, we create $O(k^2)$ variables.
At the end we create our final guess.
\begin{enumerate}[G-1.]
% counter for guesses continues
\setcounter{enumi}{\value{guesscounter}}
%%
    \item \label{FPT:guessallPermutations}
We guess the permutation of all $O(k^2)$ variables,
together with the relation of each variable to the labels of edges incident to these not yet determined edges.
Namely, for an edge $e$ that is not yet determined, we set its value to $x_e$ and check the labels of all of its neighbors,
which are determined by some other label,
and variables of the not yet determined neighbors,
and guess if $x_e$ is smaller, equal or bigger than the labels of the edges of its neighbors.
So, for any two variables $x_e$ and $x_f$, we know if $x_e < x_f$ or $x_e = x_f$, or $x_e > x_f$,
and for any neighboring edge $g$ of $e$ we know if $x_e < \lambda(g)$ or 
$x_e = \lambda(g)$, or $x_e > \lambda(g)$.
This results in $O(k^2)! = k^{O(k^2)}$ guesses
and consequently
each of the ILP instances we created up to now is further split into $k^{O(k^2)}$ new ones.
\end{enumerate}
We have now finished creating all ILP instances.
From \cref{sec:FPT-guessing} we know the structure of all guessed paths, to which we have just added also the knowledge of permutation of all variables.
We proceed with adding constraints to each of our ILP instances.
First we add all constraints for the labels of edges that we have determined up to now.
We then continue to iterate through all pairs of vertices and start adding equality (resp.~inequality) constraints for the fastest (resp.~not necessarily fastest) temporal paths between them.

We now describe how we add constraints to a path. Whenever we say that a duration of a path gives an equality or inequality constraint, we mean the following.
Let $P=(u=v_1,v_2, \dots, v_p = v)$ be the underlying path of a fastest temporal path from $u$ to $v$,
and let $Q = (u=z_1,z_2, \dots, z_r = v)$ be the underlying path of another temporal path from $u$ to $v$.
Then we know that $d(P,\lambda) = D_{u,v}$ and $d(Q, \lambda) \geq D_{u,v}$.
Using \cref{obs:durationPwithWaitingTimes}
we create an \emph{equality constraint} for $P$
of the form 
\begin{equation}\label{eq:FPT-equalityConstraint}
    D_{u,v} = \sum _ {i=2} ^ {p-1} (\lambda (v_{i}v_{i+1}) - \lambda (v_{i-1}v_i))_\Delta + 1,
\end{equation}
and an \emph{inequality constraint} for $Q$ 
\begin{equation}\label{eq:FPT-inequalityConstraint}
    D_{u,v} \leq \sum _ {i=2} ^ {r-1} (\lambda (z_{i}z_{i+1}) - \lambda (z_{i-1}z_i))_\Delta + 1.
\end{equation}
In both cases we implicitly assume that if the difference of $(\lambda (z_{i}z_{i+1}) - \lambda (z_{i-1}z_i))$ is negative, for some $i$, we add the value $\Delta$ to it (\ie we consider the difference modulo $\Delta$), therefore we have the sign $\Delta$ around the brackets.
Note that we know if the difference of two consecutive labels is positive or negative. 
In the case when two consecutive labels are determined with respect to the same label $\lambda(e)$ the difference between them is easy to determine,
if one or both consecutive labels are not yet determined 
then we have guessed in what kind of relation they are (see guess \textcolor{lipicsGray}{\textsf{\textbf{G-\ref{FPT:guessallPermutations}}}}).
Therefore we know when $\Delta$ has to be added, which implies that \cref{eq:FPT-equalityConstraint,eq:FPT-inequalityConstraint} are calculated correctly for all paths.

>>>>

%From the construction of $Z$ (\ie by exhaustively removing vertices of degree one from~$G$), it follows that $G[Z]$ (the graph induced in $G$ by $Z$) consists of disjoint trees.  Each of these trees has a unique neighbor $u_i$ in $G'$.  Denote by $T_i$ the tree obtained by considering such a vertex $u_i$ and all the trees from $G[Z]$ that are incident to $u_i$.  We then refer to $u_i$ as the \emph{clip vertex} of the tree $T_i$.

Finally, the set of \emph{important vertices} is defined as the set $U \cup U^{\ast} \cup Z^{\ast}$.
For an illustration see \cref{fig:labelingVertices}. Note that determining sets $U$ and $U^*$ takes linear time.

@@

%Let us remember the following basic property of the problem \deltaExact.
Recall that a labeling $\lambda$ of $G$ satisfies $D$ if the duration of a fastest temporal path from vertex $v_i$ to $v_j$ equals $D_{v_i, v_j}$.
%, and no other temporal path from $v_i$ to $v_j$ is faster.
In order to find a labeling that satisfies this property we split our analysis in nine cases.
We consider the fastest temporal paths 
where the starting vertex is in one of the sets $U,V(G') \setminus U, Z$,
and similarly the destination vertex is in one of the sets $U,V(G') \setminus U, Z$.
In each of these cases, we guess the underlying path $P$ that at least one fastest temporal path from the vertex $v_i$ to $v_j$ follows, 
which results in one equality constraint for the labels on the path $P$. 
For all other temporal paths from $v_i$ to $v_j$ we know that they cannot be faster, so we introduce inequality constraints for them.
This results in producing $f(k) \cdot |D| ^ {O(1)}$ constraints. 
Note that we have to do this while keeping the total number of variables upper-bounded by some function in $k$.
%We have to guarantee only that the number of guesses of fastest temporal paths is exponential in $k$ and not in $n$.

For an easier understanding and analysis of the algorithm, we give the following definition.
\begin{definition}\label{def:segments}
    Let $U \subseteq V(G')$ be a set of vertices of interest and let $u,v \in U$.
    A path $P = (u=v_1,v_2, \dots, v_p = v)$  in graph $G'$, where all inner vertices are not in $U$, \ie $v_i \notin U$ for all $i \in \{ 2, 3, \dots, p-1\}$,
    is called a \emph{segment} from $u$ to $v$. We denote it as $S_{u,v}$.
\end{definition}
%TODO: We can have two segments among a pair of vertices in U !!!
%Among a fixed pair of vertices $u,v \in U$ there can be at most two different segments. If there are two, then one has to be of length $1$, just a single edge, and the second one of length at least $2$.
Note from \cref{def:segments} that $S_{u,v} \neq S_{v,u}$ since we consider paths to be directed. It's also worth emphasizing that $S_{v,u}$ is essentially the reversed path of $S_{u,v}$.
Furthermore, it's important to observe that a temporal path in $G'$ between two vertices of interest is either a segment or consists of a sequence of segments.
Moreover, any inner vertex $v_i$ in the segment $S_{u,v}$ ($v_i \in S_{u,v} \setminus \{u,v\}$) is part of precisely two segments: $S_{u,v}$ and $S_{v,u}$.
Given that we have at most $4k$ interesting vertices in $G'$, we can deduce the following crucial result.
\begin{corollary}\label{obs:FPT-k2segments}
    There are at most $O(k^2)$ segments in $G'$.
\end{corollary}
%TODO: one reviewer asked how we get this number? (number of paths)
%For a pair $u,v$ of vertices in $U$, there are $O(k!)=O(k^k)$ possible paths in $G'$ between them.

@@

\begin{comment}
%we never refer to this lemma
\begin{lemma}\label{lemma:FPT-unlabeldPaths-01}
    Suppose that
    $S_{u,v}, S_{w,z}$ are two segments with $ v_i \in S_{u,v}$ and $z_j \in S_{w,z}$,
    where $z_j$ is a split vertex of $v_i$ in the segment $S_{w,z}$.
    W.l.o.g. suppose that the fastest temporal path from $v_i$ to $z_j$ travels through vertices $u$ and $w$.
    Then the fastest temporal path from $v_i$ to any other vertex of $S_{w,z}$, that is closer to $w$,
    travels through the same two vertices $u$ and $w$.
    Similarly it holds for the cases when the fastest temporal path travels through $w,v$ or $z,u$ or~$z,v$.
\end{lemma}
\begin{proof}
    Let $z_\ell$ be a vertex of $S_{w,z}$, that is closer to $w$ than $z$ in the segment.
    Let us denote with $P_{v_i,z_j}$ the underlying path of the fastest temporal path from $v_i$ to $z_j$.
    Denote with $P_{v_i,z_j}^\ell$
    the subpath of the fastest temporal path from $v_i$ to $z_j$, that terminates in $z_\ell$.
    We want to show that $P_{v_i,z_j}^\ell$ is an underlying path of a fastest temporal path from $v_i$ to $z_j$.
    Let us observe the following possibilities.

@@

    Third, suppose that the fastest temporal path from $v_i$ to $z_\ell$ travels through vertices $v$ and $z$.
    Denote this path as $P^3_{v_i,z_\ell}$.
    Then the duration of the temporal path from $v_i$ to $z_j$ using the subpath of $P^3_{v_i,z_\ell}$, would be strictly smaller than the duration of $(P_{v_i,z_j},\lambda)$, which cannot be possible.
\end{proof}
\end{comment}

\begin{lemma}\label{lemma:FPT-noUndeterminedEdgeInSegment}
Let $S_{u,v}= (u=v_1,v_2, \dots, v_p = v)$ be a segment in $G$.
If $S_{u,v}$ is of length at least $5$ ($p > 5$) 
then it is impossible for an inner edge $f = v_i v_{i+1}$ from $S_{u,v} \setminus \{u,v\}$ (where $f$ is an edge that is not incident to a vertex from $U$)
to not be a part of any fastest temporal path, of length at least $2$ between vertices in $S_{u,v}$.
In other words, there must exist a pair $v_j, v_{j'} \in S_{u,v}$ \st the fastest temporal path from $v_j$ to $v_{j'}$ passes through $f$.
If $S_{u,v}$ is of length $4$ then all temporal paths of length $2$ avoid the inner edge $f$ if and only if $f$ has the same label as both of the edges incident to it, while the label of the last remaining edge is determined with respect to $\lambda(f)$.
\end{lemma}

\begin{proof}
    For an easier understanding and better readability, we present the proof for $S_{u,v}$ of fixed length $5$.
    The case where $S_{u,v}$ is longer easily follows from the presented results.

@@

    \emph{Case 1: }$f = v_2v_3$.
    Let us observe the case of the fastest temporal paths between $v_{1}$ and $v_{3}$.
    Denote with $Q = (v_{1}, v_2, v_3)$ and with $P' = (v_3,v_4,v_5,v_6)$.
    From our proposition, it follows that
    \begin{itemize}
        \item the fastest temporal path $P ^ +$ from $v_1$ to $v_3$ 
    is of the following form 
    $P^+ = v_{1}  \leadsto v_6 \rightarrow v_5 \rightarrow v_4 \rightarrow v_3$,
    and
        \item the fastest temporal path $P ^ -$ from $v_{3}$ to $v_{1}$ 
    is of the following form 
    $P^- = v_{3} \rightarrow v_{4} \rightarrow v_5 \rightarrow v_6 \leadsto v_1$.
    \end{itemize}
    It follows that 
    $
    d(v_{1}, v_{3}, P^+) \leq d(v_{1}, v_{3}, Q),
    $
    and
    $
    d(v_{1}, v_{3}, P^-) \leq d(v_{1}, v_{3}, Q)
    $. 
    Note that $d(v_{1}, v_{3}, P^+) \geq 1 + d(v_6,v_3,P')$,
    and by the definition $d(v_6,v_3,P') = 1 + (t_4 - t_5)_\Delta + (t_3 - t_4)_\Delta$,
    where $(t_i - t_j)_\Delta$ denotes the difference of two consecutive labels $t_i, t_j$ modulo $\Delta$.
    Similarly holds for $d(v_{1}, v_{3}, P^+)$.
    Summing now both of the above equations we get
    \begin{equation}
    \begin{split} \label{eq:FPT-prf-fUnlabeled}
        d(v_{1}, v_{3}, P^+) + d(v_{3}, v_{1}, P^-) &\leq 
        d(v_{1}, v_{3}, Q) + d(v_{3}, v_{1}, Q) \\
        %
        1 + d(v_6,v_3,P') + 1 + d(v_3,v_6,P') &\leq d(v_{1}, v_{3}, Q) + d(v_{3}, v_{1}, Q) \\
        %
        3 + (t_4 - t_5)_\Delta + (t_3 - t_4)_\Delta +
        1 + (t_4 - t_3)_\Delta + (t_5 - t_4)_\Delta 
        &\leq 
        1 + (t_2 - t_1)_\Delta + 
        1 + (t_1 - t_2)_\Delta\\
        %
        (t_4 - t_5)_\Delta + (t_5 - t_4)_\Delta +
        (t_4 - t_3)_\Delta + (t_3 - t_4)_\Delta  + 2 
        &\leq 
        (t_2 - t_1)_\Delta + (t_1 - t_2)_\Delta.
    \end{split}
    \end{equation}
    Note that if $t_i \neq t_j$ we get that 
    the sum
    $(t_i - t_j)_\Delta + (t_j - t_i)_\Delta$ equals exactly $\Delta$,
    and if $t_i = t_j$ the sum equals $2\Delta$.
    This follows from the definition of travel delays at vertices (see \cref{obs:travel-delays-both-directions}).
    Therefore we get from \cref{eq:FPT-prf-fUnlabeled}, 
    that the right part is at most $2 \Delta$, while the left part is at least $2 \Delta + 1$,
    for any relation of labels $t_1,t_2, \dots, t_5$,
    which is a contradiction.

@@

    Let us observe now the case when $S_{u,v} = (u=v_1,v_2, v_3, v_4, v_5=v)$ is of length $4$.
    Let $f = v_2 v_3$ (the case with $f = v_3 v_4$ is symmetrical).
    Suppose that the fastest temporal paths between $v_1$ and $v_3$ do not use the edge $f$.
    We denote with $R^+$ the fastest path from $v_1$ to $v_3$, 
    which is of the form $u \leadsto v \rightarrow v_4 \rightarrow v_3$,
    and similarly 
    with $R^-$ the fastest path from $v_3$ to $v_1$, which is
    of the form $v_3 \rightarrow v_4 \rightarrow v \leadsto u$.
    We denote with $R' = (v_3, v_4, v_5)$ and with $S = (v_1,v_2,v_3)$.
    Again we get the following.
    \begin{equation*}
    \begin{split}
        d(v_{1}, v_{3}, R^+) + d(v_{3}, v_{1}, R^-) &\leq 
        d(v_{1}, v_{3}, S) + d(v_{3}, v_{1}, S) \\
        %
        1 + d(v_5,v_3,R') + 1 + d(v_3,v_5,R') &\leq d(v_{1}, v_{3}, S) + d(v_{3}, v_{1}, S) \\
        %
        (t_3 - t_4)_\Delta + (t_4 - t_3)_\Delta + 2 
        &\leq 
        (t_2 - t_1)_\Delta + (t_1 - t_2)_\Delta.        
    \end{split}
    \end{equation*}
    The only case when the equation has a valid solution is when $t_1 = t_2$ and $t_3 \neq t_4$,
    as in this case the left hand side evaluates to $\Delta + 2$, while the right side evaluates to $2 \Delta$.
    %
    Repeating the analysis for the fastest paths between $v_2$ and $v_4$,
    we conclude that the only valid solution is when $t_2 = t_3$ and $t_1 \neq t_4$.
    %
    Altogether, we get that $f$ is not a part of any fastest path of length $2$ in $S_{u,v}$ if and only if the label of edge $f$ is the same as the labels on the edges incident to it, while the last remaining edge has a different label.
    %
    Note now that the fastest temporal path from $v_2$ to $v_4$ must first use the edge $u v_2$ and finish with the edge $v_4 v_5$, and it has to be of duration $D_{v_2, v_4}$. Using \cref{lemma:FPT-uv-Labelalledges} we determine the label of the edge $v_4v_5$ with respect to $\lambda(f)$.
\end{proof}

@@

Each specific calculation takes linear time. Since there are $O(k^2)$ segments, the whole calculation takes $O(k^4)$ time.

@@

the edges of every segment are fully determined

@@

For this, we use the fact that at least one of the temporal paths between vertices in $S_{u,v}$ has to pass through $f$, when $S_{u,v}$ is of length at least $5$ or if it is of length $4$ there is a unique labeling that allows no such path to traverse $f$ (see~\cref{lemma:FPT-noUndeterminedEdgeInSegment}).

@@

    Thus we create $O(1)$ guesses for every such segment $S_{u,v}$, and $O(k^2)$ new guesses in total, as there are at most $O(k^2)$ segments.

@@

Once we know which temporal paths pass through $f$, 

@@

Note, from \cref{lemma:FPT-noUndeterminedEdgeInSegment} it follows that the above procedure holds only for segments of length at least $4$.

@@

At this point $G'$ is a graph, where each edge $e$ has a value for its label $\lambda(e)$
that depends on (\ie is a function of) some other label $\lambda(f)$ of edge $f$,
or it depends on no other label.
We now describe how we create variables and start building our ILP instances. 
For every edge $e$ in $G'$ that is incident to a vertex of interest, we create a variable $x_e$ that can have values from $ \{1, 2, \dots, \Delta\}$.
Besides that, we create one variable for each edge that is still not yet determined on a segment.
Since each vertex of interest is incident to at most $k$ edges in $G'$, and each segment has at most one extra not yet determined edge, we create $O(k^2)$ variables.
At the end we create our final guess.
\begin{enumerate}[G-1.]
% counter for guesses continues
\setcounter{enumi}{\value{guesscounter}}
%%
    \item \label{FPT:guessallPermutations}
We guess the permutation of all $O(k^2)$ variables.
%together with the relation of each variable to the labels of edges incident to these not yet determined edges.
%Namely, for an edge $e$ that is not yet determined, we set its value to $x_e$ and check the labels of all of its neighbors, which are determined by some other label, and variables of the not yet determined neighbors, and guess if $x_e$ is smaller, equal or bigger than the labels of the edges of its neighbors.
So, for any two variables $x_e$ and $x_f$, we know if $x_e < x_f$ or $x_e = x_f$, or $x_e > x_f$.
%and for any neighboring edge $g$ of $e$ we know if $x_e < \lambda(g)$ or $x_e = \lambda(g)$, or $x_e > \lambda(g)$.
This results in $O(k^2)! = k^{O(k^2)}$ guesses
and consequently
each of the ILP instances we created up to now is further split into $k^{O(k^2)}$ new ones.
\end{enumerate}
We have now finished creating all ILP instances.
From \cref{sec:FPT-guessing} we know the structure of all guessed paths, to which we have just added also the knowledge of permutation of all variables.
We proceed with adding constraints to each of our ILP instances.
First we add all constraints for the labels of edges that we have determined up to now.
We then continue to iterate through all pairs of vertices and start adding equality (resp.~inequality) constraints for the fastest (resp.~not necessarily fastest) temporal paths between them.

We now describe how we add constraints to a path. Whenever we say that a duration of a path gives an equality or inequality constraint, we mean the following.
Let $P=(u=v_1,v_2, \dots, v_p = v)$ be the underlying path of a fastest temporal path from $u$ to $v$,
and let $Q = (u=z_1,z_2, \dots, z_r = v)$ be the underlying path of another temporal path from $u$ to $v$.
Then we know that $d(P,\lambda) = D_{u,v}$ and $d(Q, \lambda) \geq D_{u,v}$.
Using \cref{obs:durationPwithWaitingTimes}
we create an \emph{equality constraint} for $P$
of the form 
\begin{equation}\label{eq:FPT-equalityConstraint}
    D_{u,v} = \sum _ {i=2} ^ {p-1} (\lambda (v_{i}v_{i+1}) - \lambda (v_{i-1}v_i))_\Delta + 1,
\end{equation}
and an \emph{inequality constraint} for $Q$ 
\begin{equation}\label{eq:FPT-inequalityConstraint}
    D_{u,v} \leq \sum _ {i=2} ^ {r-1} (\lambda (z_{i}z_{i+1}) - \lambda (z_{i-1}z_i))_\Delta + 1.
\end{equation}
In both cases we implicitly assume that if the difference of $(\lambda (z_{i}z_{i+1}) - \lambda (z_{i-1}z_i))$ is negative, for some $i$, we add the value $\Delta$ to it (\ie we consider the difference modulo $\Delta$), therefore we have the sign $\Delta$ around the brackets.
Note that we can determine if the difference of two consecutive labels is positive or negative. 
In the case when two consecutive labels are determined with respect to the same label $\lambda(e)$ the difference between them is easy to determine.
If consecutive labels are not determined with respect to the same label, both labels are considered undetermined and are assigned a variable for which we know in what kind of relation they are (see guess \textcolor{lipicsGray}{\textsf{\textbf{G-\ref{FPT:guessallPermutations}}}}).
Therefore, we know when $\Delta$ has to be added, which implies that \cref{eq:FPT-equalityConstraint,eq:FPT-inequalityConstraint} are calculated correctly for all paths.
